Title

### 2016-06-10 paper/sections/CaseStudy.tex
Some text

To simplify client code, we defined wrappers around \texttt{of} methods
for conveniently constructing circuits. For instance, the wrapper for
\texttt{Fan} is defined as follows:
\begin{lstlisting}
static Fan fan(int n) {
  return Fan.of(n);
}
\end{lstlisting}
The \texttt{circuit} object supports all the methods stated above.
Additionally, we introduce a \texttt{draw} method which renders a
circuit using Java Swing.
Calling \texttt{circuit.draw()} on the \texttt{circuit} will
display what Figure~\ref{sec:} shows.

@@

We can already draw simple pictures based on existing definitions.
For example, Figure~\ref{woman} (with auxiliary wrappers defined) shows
a woman in a red dress and blue stockings.
To actually visualize the picture as shown in the right hand side of
Figure~\ref{woman}, we give semantics to the language for accomplishing this goal.

\begin{figure*}
  \begin{tabular}{lr}
\begin{minipage}[t]{.9\textwidth}
\lstinputlisting[linerange=530-537]{../src/diagrams/Diagrams.java}%APPLY:WOMAN
\end{minipage}
&
\begin{minipage}[t]{.1\textwidth}
  \vspace{0.2em}
\includegraphics[width=1.3cm]{woman.png}
\end{minipage} \\
\end{tabular}
\caption{A woman drawn in \emph{diagrams}}
\label{woman}
\end{figure*}

\subsubsection{Animation}
We use scalable vector graphics (SVG) as the backend for rendering
pictures.
SVG aligns styled shapes by center, which requires us to determine a center
point of a picture as the origin of coordinate, and calculate
positions of all the shapes inside the picture accordingly.
To represent a position in the coordinates we define \texttt{Pos}:
\begin{lstlisting}
@Obj interface Pos {
    double _x(); double _y();
}
\end{lstlisting}
The position of a shape can then be determined by a pair of \texttt{Pos}, called
an extent,
locating the lower left and upper right conners:
\begin{lstlisting}
@Obj interface Extent {
    Pos _p1(); Pos _p2();
}
\end{lstlisting}
Their definitions are stable so we put them out of the family and annotated
with \texttt{@Obj}.

@@

\lstinputlisting[linerange=300-319]{../src/diagrams/Diagrams.java}%APPLY:TRANSFORM

\paragraph{Drawing}
To easily translate a \texttt{Picture} into SVG format, we flatten the recursively structured pictures into a non-empty
sequence of transformed styled shapes.
\texttt{Drawing} contains three lists of the same length carrying all the
information needed for generating a SVG file:

@@

% Step 1: Picture to Drawing
To generate a \texttt{Drawing} from a \texttt{Picture}, we extend the
\texttt{Picture} hierarchy with a \texttt{draw} method:
\lstinputlisting[linerange=268-270]{../src/diagrams/Diagrams.java}%APPLY:DRAW
For Place,

@@

% Step 2: Drawing to XML
SVG is based on XML and the structure of an XML file can be captured by following definitions:
\begin{lstlisting}
@Obj interface XML {
    String _tag();
    List<Attr> _attrs();
    List<XML> _xmls();
}
@Obj interface Attr {
    String _name();
    String _value();
}
\end{lstlisting}
We then define \texttt{toXML} method both in \texttt{Drawing} and \texttt{Shape}.
\texttt{Drawing} builds up a complete XML by calling the \texttt{toXML} on each
\texttt{Shape}.

The final step is to pretty print the content of an \texttt{XML} object and
write it to an file with extension
\emph{.svg}.
Then open the file using a modern browser, you will see the right side of
Figure~\ref{woman} appear on the screen.

\subsubsection{Discussion}
\cite{} contains exercises about extensions on the language, e.g.
adding new shapes. The deep embedding approach in Haskell requires modifications on the ADTs
whereas our approach can modularly introduce new language constructs.
Each sub-language retains independent extensibility.
For example, we can add a \texttt{show} method for the purpose of debugging.

>>>>

To simplify client code, we defined wrappers around \texttt{of} methods
for conveniently constructing circuits. For instance, the wrapper for
\texttt{Fan} is defined as follows:
\begin{lstlisting}
static Fan fan(int n) {
  return Fan.of(n);
}
\end{lstlisting}
The \texttt{circuit} object supports all the methods stated above.
Additionally, we introduce a \texttt{draw} method which renders a
circuit using Java Swing.
Calling \texttt{draw} on the \texttt{circuit} will
display what Figure~\ref{sec:} shows.

@@

We can already draw simple pictures based on existing definitions.
For example, Figure~\ref{woman} (with auxiliary wrappers defined) shows
a woman in a red dress and blue stockings.
To actually visualize the picture as shown in the right hand side of
Figure~\ref{woman}, we need to give semantics to the language for animation.

\begin{figure*}
  \begin{tabular}{lr}
\begin{minipage}[t]{.9\textwidth}
\lstinputlisting[linerange=537-544]{../src/diagrams/Diagrams.java}%APPLY:WOMAN
\end{minipage}
&
\begin{minipage}[t]{.1\textwidth}
  \vspace{0.2em}
\includegraphics[width=1.3cm]{woman.png}
\end{minipage} \\
\end{tabular}
\caption{A woman drawn in \emph{diagrams}}
\label{woman}
\end{figure*}

\subsubsection{Animation}
We use scalable vector graphics (SVG) as the backend for rendering
pictures.
SVG aligns styled shapes by center, which requires us to determine a center
point of a picture as the origin of coordinate, and calculate
positions of all the shapes inside the picture accordingly.
To represent a position in the coordinates we define \texttt{Pos}:
\begin{lstlisting}
@Obj interface Pos {
    double _x(); double _y();
}
\end{lstlisting}
The position of a shape can then be determined by a pair of \texttt{Pos}, called
an extent,
locating the lower left and upper right conners:
\begin{lstlisting}
@Obj interface Extent {
    Pos _p1(); Pos _p2();
}
\end{lstlisting}
These definitions are stable so we put them out of the family and annotated
with \texttt{@Obj}.

@@

\lstinputlisting[linerange=164-171]{../src/diagrams/Diagrams.java}%APPLY:TRANSFORM

\paragraph{Drawing}
To easily translate a \texttt{Picture} into SVG format, we flatten the recursively structured pictures into a non-empty
sequence of transformed styled shapes, represented by \texttt{Drawing}:

@@

% Step 1: Picture to Drawing
To generate a \texttt{Drawing} from a \texttt{Picture}, we extend the
\texttt{Picture} hierarchy with a \texttt{draw} method:
\lstinputlisting[linerange=279-281]{../src/diagrams/Diagrams.java}%APPLY:DRAW

@@

% Step 2: Drawing to XML
With all the necessary information contained in drawing, we can now generate the
SVG file. SVG is based on XML and the structure of an XML file can be captured by following definitions:
\begin{lstlisting}
@Obj interface XML {
    String _tag();
    List<Attr> _attrs();
    List<XML> _xmls();
}
@Obj interface Attr {
    String _name(); String _value();
}
\end{lstlisting}
We then extend a \texttt{toXML} method both in \texttt{Drawing} and \texttt{Shape}.
\begin{lstlisting}
interface Shape {
  XML toXML(List<Attr> styleAttrs, Transform trans);
}
interface Drawing {
  XML toXML();
}
\end{lstlisting}
Calling \texttt{toXML} on a \texttt{Drawing} will builds up a complete XML by calling the \texttt{toXML} on each \texttt{Shape}.

The final step is to pretty print the content of an \texttt{XML} object and
save it to an file with the extension \emph{svg}.
Then open the file with a modern browser, you will see the right side of
Figure~\ref{woman} appear on the screen.

\subsubsection{Discussion}
Each sub-language retains independent extensibility.
\cite{} contains exercises about extensions on the language, e.g.
adding new shapes.
The deep embedding approach in Haskell requires modifications on the ADTs
whereas our approach can modularly introduce new language constructs.
At the same time, new interpretations can be added modularly.
For example, we can add a \texttt{show} method for the purpose of debugging.

### 2016-06-10 paper/sections/Overview.tex
Some text

%As language evolves, the need of new syntax often arises.
As DSLs evolve along the time, the demand for new syntax and new semantics may arise.
It would be good if we can introduce these new features to the DSL \emph{modularly}.
%This is actually a hard problem, known as the Expression Problem (EP)~\cite{wadler}, which
%This, however, requires the host language equipped with two dimensions of extensibility.
For shallow embedding, adding new syntax is trivial - just adding new cases in
the semantic function. On the other hand, adding new semantics becomes hard as it allows
only one definition for each case. Gibbons and Wu summarized how to define
various type of semantic functions (interpretations) in shallow embedding by
applying \emph{fold theory}. However, the encoding is based on tuples, which is
not modular extensible. Though they also presented modular solution based on
datatype a lar carte\cite{}. But such solution changes the encoding style
dramatically and may be too complex to use practically.
In contrast, the OO approach allows all kinds of
interpretations to be defined in an modular intuitive way.
In this section, we will show how various types of interpretations can be
defined using the OO way with the help of \name. Along the way,
some features of \name will be introduced, which simplify our implementation significantly.

@@

There are three main changes:
1) All the original implementations are moved inside an interface \texttt{Init}; and
2) The constructor classes are now interfaces for the purpose of employing
multiple interface inheritance of Java;
3) Fields of the classes become getter methods (starting with an underscore) so that their types can be refined for retaining extensibility~\cite{}.

@@

\subsection{Multiple Interpretations}
Defining additional interpretations is not so trivial for shallow embedding,
especially for functional languages.
Suppose that we want to have a function that checks whether a circuit is constructed correctly.

Here comes the solution from Gibbons and Wu:
\lstinputlisting[language=haskell,linerange=2-8]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
where tuples are used to aggregate multiple interpretations in one definition.
The way to introduce \texttt{desugar} in Section~\ref{sec:oo} is similar to this
approach, where we use records instead.
However, neither tuples nor records are extensible - whenever we need a new
interpretation, we have to revise the original code, either expanding the arith of the
tuple or inserting a new field in the record.

@@

\subsection{Dependent Interpretations}
Interpretations which depend on other interpretations are called \emph{dependent
  interpretation}. %They are hard to express modularly in traditional shallow embedding.
Before we introduce an example of such interpretations, we first add two new
constructs to our language:
\emph{stretch ns c} inserts additional wires into the circuit \emph{c} by
summing up \emph{ns} and $above\ c_1\ c_2$ combines two circuits of the same width vertically.
Their types are given below:
\lstinputlisting[language=haskell,linerange=40-41]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES

%\lstinputlisting[linerange=56-69]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
%With these new constructs, more complex circuits can be constructed.
%For example, Figure~\ref{} shows the circuit constructed by \emph{stretch
%  [2,2,2] (fan 3) `beside` fan 1}.

Adding new constructs in \name is easy: just we declare new class-like
interfaces in an extended family:
\lstinputlisting[linerange=56-69]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
Only definitions of the new constructs \texttt{Above} and \texttt{Stretch} are given.
Their peer constructs and \texttt{Circuit} are implicitly re-declared by \name,
e.g. \texttt{Fan}:

@@

Not arbitrary circuits can be combined using \texttt{Above} and \texttt{Stretch}.
We hence define the \texttt{wellSized} method for these two new constructs to
verify the constraints they imply:
\lstinputlisting[linerange=73-85]{../src/paper/sec3/Circuit.java}%APPLY:DEPENDENT

Definitions of \texttt{wellSized} for the extended cases make it a dependent interpretation, as they
rely on another method defined on the \texttt{Circuit}, \texttt{width}, for calculating the width of a circuit.

Here, we show yet another appealing feature of \name - interpretation combination.
That is, we can modularly merge interpretations defined for the same data structure together.
The \texttt{Dependent} family combines two base families

@@

\lstinputlisting[linerange=107-139]{../src/paper/sec3/Circuit.java}%APPLY:CONTEXT_SENSITIVE

@@

\lstinputlisting[linerange=170-172]{../src/paper/sec3/Circuit.java}%APPLY:MERGE
Simply defining a new family \texttt{Merge} that extends other families we want
to combine. \name will implicitly re-declare all the interfaces and inject the
inheritance relationships. For example, the implicit redeclaration of
\texttt{Circuit} looks like this:

@@

\lstinputlisting[linerange=56-69]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX

@@

\lstinputlisting[linerange=177-189]{../src/paper/sec3/Circuit.java}%APPLY:INSTRUMENT
% TODO: not implemented yet
Family polymorphism on only gives more safety on the client code but also helps
language implementers catch bug early.
For example, if one forgets to implement, say cases from \texttt{Family}, in
\texttt{Semantics}, she would get warned they have not implemented the
\texttt{wellSized} method because class-like interfaces should not contain any
unimplemented methods except for getters.
These errors are captured by automatically re-declaring members with inheritance relationships.
%The \texttt{of} method is  returns an instance of an anonymous class
%that implements the interface with all getters implemented and its instance is returned.
\end{comment}

>>>>

%As language evolves, the need of new syntax often arises.
As DSLs evolve along the time, the demand for new syntax and new semantics may arise.
It would be good if we can introduce these new features to the DSL \emph{modularly}.
%This is actually a hard problem, known as the Expression Problem (EP)~\cite{wadler}, which
%This, however, requires the host language equipped with two dimensions of extensibility.
For shallow embedding, adding new syntax is trivial - just adding new cases in
the semantic function. On the other hand, adding new semantics becomes hard as it allows
only one definition for each case. Gibbons and Wu summarized how to define
various type of semantic functions (interpretations) in shallow embedding by
applying \emph{fold theory}. However, the encoding is based on tuples, which is
not modular extensible. Though they also presented modular solution based on
\emph{data types a la carte}\cite{swierstra2008data}. But such solution changes the encoding style
dramatically and may be too complex to use practically.
In contrast, the OO approach allows all kinds of
interpretations to be defined in an modular intuitive way.
In this section, we will show how various types of interpretations can be
defined using the OO way with the help of \name. Along the way,
some features of \name will be introduced, which simplify our implementation significantly.

@@

There are three main changes:
1) All the original implementations are moved inside an interface \texttt{Init}; and
2) The constructor classes are now interfaces for the purpose of employing
multiple interface inheritance of Java;
3) Fields of the classes become getter methods (starting with an underscore) so that their types can be refined for retaining extensibility~\cite{eptrivially16}.

@@

\subsection{Multiple Interpretations}
Defining additional interpretations is not so trivial for shallow embedding,
especially for functional languages.
Suppose that we want to have a function that checks whether a circuit is constructed correctly. Here comes the solution from Gibbons and Wu:
\lstinputlisting[language=haskell,linerange=2-8]{./code/NewSemantics.hs}%APPLY:SEMANTICS_HS
Tuples are used to aggregate multiple interpretations in one definition.
The way to introduce \texttt{desugar} demonstrated in Section~\ref{sec:oo} is similar to this
approach, where records are used instead.
However, neither tuples nor records are extensible - whenever we need a new
interpretation, we have to revise the original code, either expanding the arith of the
tuple or inserting a new field in the record.

@@

\subsection{Dependent Interpretations}
Interpretations which depend on other interpretations are called \emph{dependent
  interpretation}. %They are hard to express modularly in traditional shallow embedding.
Before we introduce an example of such interpretations, we first show how to add new
constructs in \name.

The new constructs are: \emph{stretch ns c} which inserts additional wires into the circuit \emph{c} by
summing up \emph{ns} and $above\ c_1\ c_2$ which combines two circuits of the same width vertically.
Their types are given below:
\lstinputlisting[language=haskell,linerange=40-41]{./code/shallowCircuit.hs}%APPLY:SYNTAX_TYPES

%\lstinputlisting[linerange=56-68]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
%With these new constructs, more complex circuits can be constructed.
%For example, Figure~\ref{} shows the circuit constructed by \emph{stretch
%  [2,2,2] (fan 3) `beside` fan 1}.

Adding new constructs in \name is easy - just we define new class-like
interfaces in an extended family:
\lstinputlisting[linerange=56-68]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX
Only definitions of the new constructs \texttt{Above} and \texttt{Stretch} are given.
Their peer constructs and \texttt{Circuit} are implicitly re-declared by \name,
e.g. \texttt{Fan}:

@@

Not arbitrary circuits can be combined using \texttt{Above} and \texttt{Stretch}
as stated in their specifications.
We hence define the \texttt{wellSized} method for these two new constructs to
verify the constraints they imply:
\lstinputlisting[linerange=72-84]{../src/paper/sec3/Circuit.java}%APPLY:DEPENDENT

Definitions of \texttt{wellSized} for the extended cases make it a dependent interpretation, as they
rely on another method defined on the \texttt{Circuit}, \texttt{width}, for calculating the width of a circuit.
Most importantly, our approach do not force dependent interpretation to be
defined together with what they depend on. In the case of \texttt{width} and
\texttt{wellSized}, they can be defined in separate families as long as the
signature of \texttt{width} is exposed to the definition of \texttt{wellSized}.

The definition of \texttt{Dependent} shows yet another appealing feature of \name - interpretation combination.
That is, we can modularly merge interpretations defined for the same data structure together.
The \texttt{Dependent} family combines two base families

@@

\lstinputlisting[linerange=106-138]{../src/paper/sec3/Circuit.java}%APPLY:CONTEXT_SENSITIVE

@@

\lstinputlisting[linerange=169-171]{../src/paper/sec3/Circuit.java}%APPLY:MERGE
Simply defining a new family \texttt{Merge} that extends other families we want
to combine. \name will implicitly re-declare all the interfaces and inject the
inheritance relationships. For example, the implicit redeclaration of
\texttt{Circuit} looks like this:

@@

\lstinputlisting[linerange=56-68]{../src/paper/sec3/Circuit.java}%APPLY:SYNTAX

@@

\lstinputlisting[linerange=176-188]{../src/paper/sec3/Circuit.java}%APPLY:INSTRUMENT
% TODO: not implemented yet
Family polymorphism on only gives more safety on the client code but also helps
language implementers catch bug early.
For example, if one forgets to implement, say cases from \texttt{Family}, in
\texttt{Semantics}, she would get warned they have not implemented the
\texttt{wellSized} method because class-like interfaces should not contain any
unimplemented methods except for getters.
These errors are captured by automatically re-declaring members with inheritance relationships.
%The \texttt{of} method is  returns an instance of an anonymous class
%that implements the interface with all getters implemented and its instance is returned.
\end{comment}

### 2016-06-10 paper/sections/RelatedWork.tex
Some text

In OOP, the Vistor pattern is widely used for modeling languages, which is
well-known for adding new operations easily. Many works
have been done for making it possible to extend variants~\cite{oliveira07genericity,oliveira09modular}.
%In fact, the Visitor pattern has a strong connection with deep and shallow embedding.
% Specifically, the internal variant corresponds to shallow embedding whereas the external variant is equivalent to deep embedding.
Object Algebras~\cite{bruno12oa} simplifies the solution, which is an variant of
internal visitor pattern.
 % However, this approach requires heavy encoding and advanced type system features in Scala.
Another line of work is on OO decomposition. ~\cite{zenger} capture recursive
arguments using virtual types with bounds. Wang and Oliveira~\cite{eptrivially16}
uses covariant types, a common feature available in most OO languages, to achieve the same goal which is available in most OO languages.
Compared to the Visitor pattern, the OO decomposition approach is simpler and our encoding is based on ~\cite{eptrivially16}.

In FP, the two main solutions are \emph{finally
  tagless}~\cite{carette2009finally} and \emph{data types a la
  carte}~\cite{swierstra2008data} (DTC).
Finally tagless approach uses a type class to abstract over all interpretations
of the language. Concrete interpretations are given through creating a data type and
making it an instance of that type.
DTC represents language constructs separately and composes them together using
extensible sums. However, not like OO languages which come with subtyping, one
has to manually implement the subtyping machinery for variants.
Moreover, neither DTC nor finally tagless approach can add new interpretations
that depend on existing interpretations modularly like our approach.

>>>>

In OOP, the Vistor pattern is widely used for modeling languages, which is
well-known for adding new operations easily. Many works
have been done for making it possible to extend variants~\cite{oliveira07genericity,oliveira09modular}.
%In fact, the Visitor pattern has a strong connection with deep and shallow embedding.
% Specifically, the internal variant corresponds to shallow embedding whereas the external variant is equivalent to deep embedding.
Object Algebras~\cite{bruno12oa} simplifies the solution, which is an variant of
internal visitor pattern.
 % However, this approach requires heavy encoding and advanced type system features in Scala.
Another line of work is on OO decomposition. ~\cite{zenger} capture recursive
arguments using virtual types with bounds using Scala. Wang and Oliveira~\cite{eptrivially16}
uses covariant types, a common feature available in most OO languages, to
achieve same goal, which is the basis of our approach.
Compared to the Visitor pattern, the OO decomposition approach is simpler and
our encoding is based on ~\cite{eptrivially16}. The
naming problem for extensions is solved by family polymorphism.

In FP, the two main solutions are \emph{finally
  tagless}~\cite{carette2009finally} and \emph{data types a la
  carte}~\cite{swierstra2008data} (DTC).
Finally tagless approach uses a type class to abstract over all interpretations
of the language. Concrete interpretations are given through creating a data type and
making it an instance of that type.
DTC represents language constructs separately and composes them together using
extensible sums. However, not like OO languages which come with subtyping, one
has to manually implement the subtyping machinery for variants.
Moreover, both DTC and finally tagless approach force dependent interpretations
to be defined along with what they depend on.
This prevents new interpretation that depend on existing interpretations from
being defined modularly while our approach does not have such requirement.
